---
title: "P8105 HW3"
author: "Veerapetch Petchger"
date: "2025-10-05"
output: github_document
---
## Problem 1
```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations. Then, do or answer the following (commenting on the results of each):

1. How many aisles are there, and which aisles are the most items ordered from?
```{r most_items}
instacart %>% 
  count(aisle, name = "n_obs") %>%
  arrange(desc(n_obs)) %>% 
  head()
```

There are `r instacart %>% summarise(n_aisles = n_distinct(aisle))` different aisles. The `fresh vegetables` and `fresh fruits` are by far the most popular aisles, followed by `packaged vegetables fruits`, `yogurt`, and `packaged cheese`.

2. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r 10000}
instacart %>%
  count(aisle, name = "n_obs") %>%
  filter(n_obs > 10000) %>%
  mutate(aisle = forcats::fct_reorder(aisle, n_obs)) %>%
  ggplot(aes(x = aisle, y = n_obs, color = aisle)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Aisles With Over 10000 Items Ordered",
    x = "Aisle Names",
    y = "Number of of Items Ordered"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "none")
```

3. Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r top_3}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(order_count = n()) %>%
  arrange(aisle, desc(order_count)) %>%
  slice_max(order_count, n = 3)
```

4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r mean_hour}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour,
    names_prefix = "Day_"
  )
```

## Problem 2
This Problem uses the Zillow datasets introduced in Homework 2. Both datasets are available here. Import, clean, and otherwise tidy these datasets.
```{r zillow}
zip_codes_df =
  read_csv("zillow_data/Zip Codes.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    zip_code = as.character(zip_code),
    borough  = recode(
      county,
      "New York" = "Manhattan",
      "Kings"    = "Brooklyn",
      "Queens"   = "Queens",
      "Bronx"    = "Bronx",
      "Richmond" = "Staten Island",
    )
  ) %>%
  select(zip_code, borough)

zora_df =
  read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    cols = x2015_01_31:x2024_08_31,
    names_to = "ym_str",
    values_to = "rent"
  ) %>%
  rename(zip_code = region_name) %>% 
  mutate(
    ym_str = stringr::str_remove(ym_str, "^x"),
    ym_str = stringr::str_replace_all(ym_str, "_", "-"),
    date   = lubridate::ymd(ym_str),
    year   = lubridate::year(date),
    month  = lubridate::month(date),
    day    = lubridate::day(date),
    zip_code = as.character(zip_code))

zillow_df =
  zora_df %>% 
  left_join(zip_codes_df, by = "zip_code")
```

There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?

Create a reader-friendly table showing the average rental price in each borough and year (not month). Comment on trends in this table.

Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.

Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot.

Combine the two previous plots into a single graphic, and export this to a results folder in your repository.


